Very simple spline models

### 2024-07-06 text/21/glm-diagnostics.Rmd
Updated 2022 Rmd files

---
title: "Diagnostics for a generalized linear model"
author: "Steve Simon"
source: new
date: "2021-10-18"
categories:
- Blog post
tags:
- Generalized linear model
output: html_document
---

@@

```{r setup}
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(foreign)))
suppressMessages(suppressWarnings(library(magrittr)))
suppressMessages(suppressWarnings(library(pscl)))
```

@@

```{r, eval=FALSE}
fn <- "http://faculty.missouri.edu/huangf/data/poisson/articles.csv"
articles <- read.csv(fn)
head(articles)
```

@@

```{r}
fn <- "http://www.pmean.com/00files/couart2.dta"
articles <- read.dta(fn)
head(articles)
```

@@

```{r}
ggplot(articles, aes(art)) + 
  geom_histogram(binwidth=1, color="black", fill="white")
```

@@

```{r}
data.frame(x=0:max(articles$art)) %>%
  mutate(p=dpois(x, lambda=mean(articles$art))) %>%
  ggplot(aes(factor(x), p)) +
    geom_col(color="black", fill="white")
```

@@

```{r}
articles %>%
  group_by(fem) %>%
  summarize(n=n())
```

```{r}
quantile(articles$ment)
```

```{r}
quantile(articles$phd)
```

```{r}
articles %>%
  group_by(mar) %>%
  summarize(n=n())
```

```{r}
articles %>%
  group_by(kid5) %>%
  summarize(n=n())
```

@@

```{r}
articles %>%
  mutate(ment=(ment-min(ment))/(max(ment)-min(ment))) %>%
  mutate(phd=(phd-min(phd))/(max(phd)-min(phd))) -> articles_scaled
```

@@

```{r}
ggplot(articles, aes(factor(fem), art)) +
  geom_boxplot()
```

```{r}
ggplot(articles, aes(ment, art)) +
  geom_point() + 
  geom_smooth()
```

```{r}
ggplot(articles, aes(phd, art)) +
  geom_point() + 
  geom_smooth()
```

```{r}
ggplot(articles, aes(factor(mar), art)) +
  geom_boxplot()
```

```{r}
ggplot(articles, aes(factor(kid5), art)) +
  geom_boxplot()
```

@@

```{r}
pois_fem <- glm(art ~ fem, family=poisson, data = articles_scaled)
summary(pois_fem)
```

```{r}
pois_ment <- glm(art ~ ment, family=poisson, data = articles_scaled)
summary(pois_ment)
```

```{r}
pois_phd <- glm(art ~ phd, family=poisson, data = articles_scaled)
summary(pois_phd)
```

```{r}
pois_mar <- glm(art ~ mar, family=poisson, data = articles_scaled)
summary(pois_mar)
```

```{r}
pois_kid5 <- glm(art ~ kid5, family=poisson, data = articles_scaled)
summary(pois_kid5)
```

@@

```{r}
pois_all_ivs <- glm(art ~ fem + ment + phd + mar + kid5, family=poisson, data = articles_scaled)
summary(pois_all_ivs)
```

```{r}
tidy(pois_fem) %>%
  bind_rows(tidy(pois_ment)) %>%
  bind_rows(tidy(pois_phd)) %>%
  bind_rows(tidy(pois_mar)) %>%
  bind_rows(tidy(pois_kid5)) %>%
  filter(term != "(Intercept)") %>%
  mutate(unadjusted_rate_ratio=round(exp(estimate), 2)) %>%
  select(term, unadjusted_rate_ratio) -> pois_crude

@@

```{r}
r_pois <- augment(pois_all_ivs, articles_scaled, type.predict="response", type.residuals="deviance")
ggplot(r_pois, aes(.fitted, .resid)) +
  geom_point() +
  geom_smooth()
```

@@

```{r}
r_pois %>%
  filter(.fitted > 7) %>%
  data.frame
```

@@

```{r}
ggplot(r_pois, aes(factor(fem), .resid)) + 
  geom_boxplot()
```

```{r}
ggplot(r_pois, aes(ment, .resid)) + 
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(r_pois, aes(phd, .resid)) + 
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(r_pois, aes(factor(mar), .resid)) + 
  geom_boxplot()
```

```{r}
ggplot(r_pois, aes(factor(kid5), .resid)) + 
  geom_boxplot()
```

@@

```{r}
articles %>%
  mutate(ment=pmin(40, ment)) %>% 
  mutate(ment=(ment-min(ment))/(max(ment)-min(ment))) %>%
  mutate(phd=(phd-min(phd))/(max(phd)-min(phd))) -> articles_rescaled
pois_rescaled <- 
  glm(
    art ~ fem + ment + phd + mar + kid5, 
      family=poisson, 
        data = articles_rescaled)
r_rescaled <- 
  augment(
    pois_rescaled, 
      articles_rescaled, 
        type.predict="response", 
          type.residuals="deviance") 
ggplot(r_rescaled, aes(.fitted, .resid)) +
  geom_point() +
  geom_smooth()
ggplot(r_rescaled, aes(ment, .resid)) +
  geom_point() +
  geom_smooth()
```

@@

```{r}
library(MASS)
nb1 <- glm.nb(art ~ fem + ment + phd + mar + kid5, data = articles_rescaled)
summary(pois_rescaled)
summary(nb1)
```

@@

```{r}
zip1 <- 
  zeroinfl(
    art ~ fem + ment + phd + mar + kid5 | 
      1, 
        dist="poisson", 
          data = articles_rescaled)
summary(zip1)
```

```{r}
zip2 <- 
  zeroinfl(
    art ~ fem + ment + phd + mar + kid5 | 
      fem + ment + phd + mar + kid5, 
        dist="poisson",
          data = articles_rescaled)
summary(zip2)
```

```{r}
zip3 <- 
  zeroinfl(
    art ~ fem + ment + phd + mar + kid5 | 
      ment, 
        dist="poisson",
          data = articles_rescaled)
summary(zip3)
```

```{r}
zinb <- 
  zeroinfl(
    art ~ fem + ment + phd + mar + kid5 | 
      ment, 
        dist="negbin",
          data = articles_rescaled)
summary(zinb)
```

@@

```{r}
save.image(file="glm_diagnostics.RData")
```

>>>>

---
title: "Diagnostics for a generalized linear model"
author: "Steve Simon"
source: new
date: "2021-10-18"
categories:
- Blog post
tags:
- Generalized linear model
- Incomplete page
- Rmarkdown graphs
output: html_document
page_update: complete
---

@@

```{r glm-diagnostics-01}
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(foreign)))
suppressMessages(suppressWarnings(library(magrittr)))
suppressMessages(suppressWarnings(library(pscl)))
```

@@

```{r glm-diagnostics-02, eval=FALSE}
fn <- "http://faculty.missouri.edu/huangf/data/poisson/articles.csv"
articles <- read.csv(fn)
head(articles)
```

@@

```{r glm-diagnostics-03}
fn <- "http://www.pmean.com/00files/couart2.dta"
articles <- read.dta(fn)
head(articles)
```

@@

```{r glm-diagnostics-04}
ggplot(articles, aes(art)) + 
  geom_histogram(binwidth=1, color="black", fill="white")
```

@@

```{r glm-diagnostics-05}
data.frame(x=0:max(articles$art)) %>%
  mutate(p=dpois(x, lambda=mean(articles$art))) %>%
  ggplot(aes(factor(x), p)) +
    geom_col(color="black", fill="white")
```

@@

```{r glm-diagnostics-06}
articles %>%
  group_by(fem) %>%
  summarize(n=n())
```

```{r glm-diagnostics-07}
quantile(articles$ment)
```

```{r glm-diagnostics-08}
quantile(articles$phd)
```

```{r glm-diagnostics-09}
articles %>%
  group_by(mar) %>%
  summarize(n=n())
```

```{rglm-diagnostics-10}
articles %>%
  group_by(kid5) %>%
  summarize(n=n())
```

@@

```{r glm-diagnostics-11}
articles %>%
  mutate(ment=(ment-min(ment))/(max(ment)-min(ment))) %>%
  mutate(phd=(phd-min(phd))/(max(phd)-min(phd))) -> articles_scaled
```

@@

```{r glm-diagnostics-12}
ggplot(articles, aes(factor(fem), art)) +
  geom_boxplot()
```

```{r glm-diagnostics-13}
ggplot(articles, aes(ment, art)) +
  geom_point() + 
  geom_smooth()
```

```{r glm-diagnostics-14}
ggplot(articles, aes(phd, art)) +
  geom_point() + 
  geom_smooth()
```

```{r glm-diagnostics-15}
ggplot(articles, aes(factor(mar), art)) +
  geom_boxplot()
```

```{r glm-diagnostics-16}
ggplot(articles, aes(factor(kid5), art)) +
  geom_boxplot()
```

@@

```{r glm-diagnostics-17}
pois_fem <- glm(art ~ fem, family=poisson, data = articles_scaled)
summary(pois_fem)
```

```{r glm-diagnostics-18}
pois_ment <- glm(art ~ ment, family=poisson, data = articles_scaled)
summary(pois_ment)
```

```{r glm-diagnostics-19}
pois_phd <- glm(art ~ phd, family=poisson, data = articles_scaled)
summary(pois_phd)
```

```{r glm-diagnostics-20}
pois_mar <- glm(art ~ mar, family=poisson, data = articles_scaled)
summary(pois_mar)
```

```{r glm-diagnostics-21}
pois_kid5 <- glm(art ~ kid5, family=poisson, data = articles_scaled)
summary(pois_kid5)
```

@@

```{r glm-diagnostics-22}
pois_all_ivs <- glm(art ~ fem + ment + phd + mar + kid5, family=poisson, data = articles_scaled)
summary(pois_all_ivs)
```

```{r glm-diagnostics-23}
tidy(pois_fem) %>%
  bind_rows(tidy(pois_ment)) %>%
  bind_rows(tidy(pois_phd)) %>%
  bind_rows(tidy(pois_mar)) %>%
  bind_rows(tidy(pois_kid5)) %>%
  filter(term != "(Intercept)") %>%
  mutate(unadjusted_rate_ratio=round(exp(estimate), 2)) %>%
  select(term, unadjusted_rate_ratio) -> pois_crude

@@

```{r glm-diagnostics-24}
r_pois <- augment(pois_all_ivs, articles_scaled, type.predict="response", type.residuals="deviance")
ggplot(r_pois, aes(.fitted, .resid)) +
  geom_point() +
  geom_smooth()
```

@@

```{r glm-diagnostics-26}
r_pois %>%
  filter(.fitted > 7) %>%
  data.frame
```

@@

```{r glm-diagnostics-27}
ggplot(r_pois, aes(factor(fem), .resid)) + 
  geom_boxplot()
```

```{r glm-diagnostics-28}
ggplot(r_pois, aes(ment, .resid)) + 
  geom_point() +
  geom_smooth()
```

```{r glm-diagnostics-29}
ggplot(r_pois, aes(phd, .resid)) + 
  geom_point() +
  geom_smooth()
```

```{r glm-diagnostics-30}
ggplot(r_pois, aes(factor(mar), .resid)) + 
  geom_boxplot()
```

```{r glm-diagnostics-31}
ggplot(r_pois, aes(factor(kid5), .resid)) + 
  geom_boxplot()
```

@@

```{r glm-diagnostics-32}
articles %>%
  mutate(ment=pmin(40, ment)) %>% 
  mutate(ment=(ment-min(ment))/(max(ment)-min(ment))) %>%
  mutate(phd=(phd-min(phd))/(max(phd)-min(phd))) -> articles_rescaled
pois_rescaled <- 
  glm(
    art ~ fem + ment + phd + mar + kid5, 
      family=poisson, 
        data = articles_rescaled)
r_rescaled <- 
  augment(
    pois_rescaled, 
      articles_rescaled, 
        type.predict="response", 
          type.residuals="deviance") 
ggplot(r_rescaled, aes(.fitted, .resid)) +
  geom_point() +
  geom_smooth()
ggplot(r_rescaled, aes(ment, .resid)) +
  geom_point() +
  geom_smooth()
```

@@

```{r glm-diagnostics-33}
library(MASS)
nb1 <- glm.nb(art ~ fem + ment + phd + mar + kid5, data = articles_rescaled)
summary(pois_rescaled)
summary(nb1)
```

@@

```{r glm-diagnostics-34}
zip1 <- 
  zeroinfl(
    art ~ fem + ment + phd + mar + kid5 | 
      1, 
        dist="poisson", 
          data = articles_rescaled)
summary(zip1)
```

```{r glm-diagnostics-35}
zip2 <- 
  zeroinfl(
    art ~ fem + ment + phd + mar + kid5 | 
      fem + ment + phd + mar + kid5, 
        dist="poisson",
          data = articles_rescaled)
summary(zip2)
```

```{r glm-diagnostics-36}
zip3 <- 
  zeroinfl(
    art ~ fem + ment + phd + mar + kid5 | 
      ment, 
        dist="poisson",
          data = articles_rescaled)
summary(zip3)
```

```{r glm-diagnostics-37}
zinb <- 
  zeroinfl(
    art ~ fem + ment + phd + mar + kid5 | 
      ment, 
        dist="negbin",
          data = articles_rescaled)
summary(zinb)
```

@@

```{r glm-diagnostics-99}
save.image(file="glm_diagnostics.RData")
```

### 2024-07-06 text/21/poisson-inferences.Rmd
Updated 2022 Rmd files

---
title: "Working with the Poisson distribution"
author: "Steve Simon"
source: new
date: "2021-08-30"
categories:
- Blog post
tags:
- Probability distributions
output: html_document
---

@@

```{r setup}
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(magrittr)))
```

@@

```{r bar-charts}
data.frame(x=0:25) %>%
  mutate(p=dpois(x, 0.5)) %>%
  ggplot(aes(x,p)) +
    geom_col() +
    expand_limits(y=1) +
    ggtitle("Poisson probabilities with lambda=0.5")
data.frame(x=0:25) %>%
  mutate(p=dpois(x, 2)) %>%
  ggplot(aes(x,p)) +
    geom_col() +
    expand_limits(y=1) +
    ggtitle("Poisson probabilities with lambda=2")
data.frame(x=0:25) %>%
  mutate(p=dpois(x, 8)) %>%
  ggplot(aes(x,p)) +
    geom_col() +
    expand_limits(y=1) +
    ggtitle("Poisson probabilities with lambda=8")
```

@@

```{r}
rate_per_thousand <- rate*1000
rate_per_thousand
```

@@

```{r}
count_ci <- c(x-1.96*sqrt(x), x+1.96*sqrt(x))
count_ci
```

@@

```{r}
rate_ci <- count_ci * 1000 / n
rate_ci
```

@@

```{r}
round(rate_ci, 1)
```

@@

```{r}
log_count <- log(x)
log_ci <- c(log_count-1.96*sqrt(1/x), log_count+1.96*sqrt(1/x))
log_ci
```

@@

```{r}
back_transformed_ci <- exp(log_ci)
back_transformed_ci
```

@@

```{r}
rate_ci <- back_transformed_ci * 1000 / n
rate_ci
```

@@

```{r}
round(rate_ci, 1)
```

@@

```{r}
ln_ratio <- log(r1/r2)
se_ratio <- sqrt(1/x1+1/x2)
se_ratio
ci1 <- c(ln_ratio-1.96*se_ratio, ln_ratio+1.96*se_ratio)
ci1
```

@@

```{r}
ci2 <- exp(ci1)
ci2
```

@@

```{r}
injuries <- data.frame(x=c(x1, x2), T=c(T1, T2), g=1:0)
glm(x~g, family=poisson, offset=log(T), data=injuries) %>%
  tidy -> c
c
```

>>>>

---
title: "Working with the Poisson distribution"
author: "Steve Simon"
source: new
date: "2021-08-30"
categories:
- Blog post
tags:
- Incomplete page
- Probability distributions
- Rmarkdown graphs
output: html_document
page_update: complete
---

@@

```{r poisson-inferences-01}
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(magrittr)))
```

@@

```{r  poisson-inferences-02}
data.frame(x=0:25) %>%
  mutate(p=dpois(x, 0.5)) %>%
  ggplot(aes(x,p)) +
    geom_col() +
    expand_limits(y=1) +
    ggtitle("Poisson probabilities with lambda=0.5")
data.frame(x=0:25) %>%
  mutate(p=dpois(x, 2)) %>%
  ggplot(aes(x,p)) +
    geom_col() +
    expand_limits(y=1) +
    ggtitle("Poisson probabilities with lambda=2")
data.frame(x=0:25) %>%
  mutate(p=dpois(x, 8)) %>%
  ggplot(aes(x,p)) +
    geom_col() +
    expand_limits(y=1) +
    ggtitle("Poisson probabilities with lambda=8")
```

@@

```{r poisson-inferences-03}
rate_per_thousand <- rate*1000
rate_per_thousand
```

@@

```{r poisson-inferences-04}
count_ci <- c(x-1.96*sqrt(x), x+1.96*sqrt(x))
count_ci
```

@@

```{r poisson-inferences-05}
rate_ci <- count_ci * 1000 / n
rate_ci
```

@@

```{r poisson-inferences-06}
round(rate_ci, 1)
```

@@

```{r poisson-inferences-07}
log_count <- log(x)
log_ci <- c(log_count-1.96*sqrt(1/x), log_count+1.96*sqrt(1/x))
log_ci
```

@@

```{r poisson-inferences-08}
back_transformed_ci <- exp(log_ci)
back_transformed_ci
```

@@

```{r poisson-inferences-09}
rate_ci <- back_transformed_ci * 1000 / n
rate_ci
```

@@

```{r poisson-inferences-10}
round(rate_ci, 1)
```

@@

```{r poisson-inferences-11}
ln_ratio <- log(r1/r2)
se_ratio <- sqrt(1/x1+1/x2)
se_ratio
ci1 <- c(ln_ratio-1.96*se_ratio, ln_ratio+1.96*se_ratio)
ci1
```

@@

```{r poisson-inferences-12}
ci2 <- exp(ci1)
ci2
```

@@

```{r poisson-inferences-13}
injuries <- data.frame(x=c(x1, x2), T=c(T1, T2), g=1:0)
glm(x~g, family=poisson, offset=log(T), data=injuries) %>%
  tidy -> c
c
```

### 2024-07-06 text/21/poisson-regression.Rmd
Updated 2022 Rmd files

```{r setup}
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(foreign)))
suppressMessages(suppressWarnings(library(magrittr)))
```

@@

```{r, eval=FALSE}
fn <- "http://faculty.missouri.edu/huangf/data/poisson/articles.csv"
articles <- read.csv(fn)
head(articles)
```

@@

```{r}
fn <- "http://www.pmean.com/00files/couart2.dta"
articles <- read.dta(fn)
head(articles)
```

@@

```{r}
pois1 <- glm(art ~ fem + ment + phd + mar + kid5, family=poisson, data = articles)
summary(pois1)
```

@@

```{r}
pois1 %>%
  tidy %>%
  mutate(lower_limit=estimate-1.96*std.error) %>%
  mutate(upper_limit=estimate+1.96*std.error) %>%
  select(term, estimate, lower_limit, upper_limit) -> log_coefficients
log_coefficients
```

```{r}
log_coefficients %>%
  mutate(estimate=exp(estimate)) %>%
  mutate(lower_limit=exp(lower_limit)) %>%
  mutate(upper_limit=exp(upper_limit)) -> multiplicative_coefficients
multiplicative_coefficients
```

@@

```{r}
fn <- "http://www.statsci.org/data/general/motorins.txt"
swe <- read.delim(fn, header=TRUE, sep="\t")
head(swe)
```

@@

```{r}
pois2 <- 
  glm(
    Claims~offset(log(Insured)) + factor(Kilometres) + factor(Zone) + Bonus + factor(Make),
    data=swe, family=poisson)
summary(pois2)
```

@@

```{r}
pois2 %>%
  tidy %>%
  mutate(lower_limit=estimate-1.96*std.error) %>%
  mutate(upper_limit=estimate+1.96*std.error) %>%
  select(term, estimate, lower_limit, upper_limit) %>%
  mutate(estimate=exp(estimate)) %>%
  mutate(lower_limit=exp(lower_limit)) %>%
  mutate(upper_limit=exp(upper_limit)) -> multiplicative_coefficients
multiplicative_coefficients
```

@@

```{r}
mean(articles$art)
var(articles$art)
```

@@

```{r}
glance(pois1)
```

>>>>

```{r poisson-regression-01}
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(foreign)))
suppressMessages(suppressWarnings(library(magrittr)))
```

@@

```{r poisson-regression-02, eval=FALSE}
fn <- "http://faculty.missouri.edu/huangf/data/poisson/articles.csv"
articles <- read.csv(fn)
head(articles)
```

@@

```{r poisson-regression-03}
fn <- "http://www.pmean.com/00files/couart2.dta"
articles <- read.dta(fn)
head(articles)
```

@@

```{r poisson-regression-04}
pois1 <- glm(art ~ fem + ment + phd + mar + kid5, family=poisson, data = articles)
summary(pois1)
```

@@

```{r poisson-regression-05}
pois1 %>%
  tidy %>%
  mutate(lower_limit=estimate-1.96*std.error) %>%
  mutate(upper_limit=estimate+1.96*std.error) %>%
  select(term, estimate, lower_limit, upper_limit) -> log_coefficients
log_coefficients
```

```{r poisson-regression-06}
log_coefficients %>%
  mutate(estimate=exp(estimate)) %>%
  mutate(lower_limit=exp(lower_limit)) %>%
  mutate(upper_limit=exp(upper_limit)) -> multiplicative_coefficients
multiplicative_coefficients
```

@@

```{r poisson-regression-07}
fn <- "http://www.statsci.org/data/general/motorins.txt"
swe <- read.delim(fn, header=TRUE, sep="\t")
head(swe)
```

@@

```{r poisson-regression-08}
pois2 <- 
  glm(
    Claims~offset(log(Insured)) + factor(Kilometres) + factor(Zone) + Bonus + factor(Make),
    data=swe, family=poisson)
summary(pois2)
```

@@

```{r poisson-regression-09}
pois2 %>%
  tidy %>%
  mutate(lower_limit=estimate-1.96*std.error) %>%
  mutate(upper_limit=estimate+1.96*std.error) %>%
  select(term, estimate, lower_limit, upper_limit) %>%
  mutate(estimate=exp(estimate)) %>%
  mutate(lower_limit=exp(lower_limit)) %>%
  mutate(upper_limit=exp(upper_limit)) -> multiplicative_coefficients
multiplicative_coefficients
```

@@

```{r poisson-regression-10}
mean(articles$art)
var(articles$art)
```

@@

```{r poisson-regression-11}
glance(pois1)
```

### 2024-07-06 text/21/quasi-likelihood.Rmd
Updated 2022 Rmd files

```{r setup}
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(foreign)))
suppressMessages(suppressWarnings(library(magrittr)))
```

@@

```{r, eval=FALSE}
fn <- "http://faculty.missouri.edu/huangf/data/poisson/articles.csv"
articles <- read.csv(fn)
head(articles)
```

@@

```{r}
fn <- "http://www.pmean.com/00files/couart2.dta"
articles <- read.dta(fn)
head(articles)
```

@@

```{r}
pois1 <- glm(art ~ fem + ment + phd + mar + kid5, family=poisson, data = articles)
summary(pois1)
```

@@

```{r}
pois1 %>%
  tidy %>%
  mutate(lower_limit=estimate-1.96*std.error) %>%
  mutate(upper_limit=estimate+1.96*std.error) %>%
  select(term, estimate, lower_limit, upper_limit) %>%
  mutate(estimate=exp(estimate)) %>%
  mutate(lower_limit=exp(lower_limit)) %>%
  mutate(upper_limit=exp(upper_limit)) -> pois1_coefficients
pois1_coefficients
```

@@

```{r}
quasi1 <- glm(art ~ fem + ment + phd + mar + kid5, family=quasipoisson, data = articles)
summary(quasi1)
```

@@

```{r}
quasi1 %>%
  tidy %>%
  mutate(lower_limit=estimate-1.96*std.error) %>%
  mutate(upper_limit=estimate+1.96*std.error) %>%
  select(term, estimate, lower_limit, upper_limit) %>%
  mutate(estimate=exp(estimate)) %>%
  mutate(lower_limit=exp(lower_limit)) %>%
  mutate(upper_limit=exp(upper_limit)) -> quasi1_coefficients
quasi1_coefficients
```

>>>>

```{r quasi-likelihood-01}
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(foreign)))
suppressMessages(suppressWarnings(library(magrittr)))
```

@@

```{r quasi-likelihood-02, eval=FALSE}
fn <- "http://faculty.missouri.edu/huangf/data/poisson/articles.csv"
articles <- read.csv(fn)
head(articles)
```

@@

```{r quasi-likelihood-03}
fn <- "http://www.pmean.com/00files/couart2.dta"
articles <- read.dta(fn)
head(articles)
```

@@

```{r quasi-likelihood-04}
pois1 <- glm(art ~ fem + ment + phd + mar + kid5, family=poisson, data = articles)
summary(pois1)
```

@@

```{r quasi-likelihood-05}
pois1 %>%
  tidy %>%
  mutate(lower_limit=estimate-1.96*std.error) %>%
  mutate(upper_limit=estimate+1.96*std.error) %>%
  select(term, estimate, lower_limit, upper_limit) %>%
  mutate(estimate=exp(estimate)) %>%
  mutate(lower_limit=exp(lower_limit)) %>%
  mutate(upper_limit=exp(upper_limit)) -> pois1_coefficients
pois1_coefficients
```

@@

```{r quasi-likelihood-06}
quasi1 <- glm(art ~ fem + ment + phd + mar + kid5, family=quasipoisson, data = articles)
summary(quasi1)
```

@@

```{r quasi-likelihood-07}
quasi1 %>%
  tidy %>%
  mutate(lower_limit=estimate-1.96*std.error) %>%
  mutate(upper_limit=estimate+1.96*std.error) %>%
  select(term, estimate, lower_limit, upper_limit) %>%
  mutate(estimate=exp(estimate)) %>%
  mutate(lower_limit=exp(lower_limit)) %>%
  mutate(upper_limit=exp(upper_limit)) -> quasi1_coefficients
quasi1_coefficients
```

### 2024-07-06 text/21/read-acs-data.Rmd
Updated 2022 Rmd files

---
title: "Reading information from the American Community Survey into R"
source: new
date: "2021-09-29"
categories:
- Recommendation
tags:
- Website details
output: html_document
---

@@

```{r}
library(tidycensus)
library(tidyverse)
library(viridis)
key <- Sys.getenv("CENSUS_API_KEY")
census_api_key("11bafd82737f7d75b3305ada1e772501d0eab002")
tarr <- get_acs(geography = "tract", variables = "B19013_001",
state = "TX", county = "Tarrant", geometry = TRUE)
ggplot(tarr, aes(fill = estimate, color = estimate)) +
geom_sf() +
coord_sf(crs = 26914) +
scale_fill_viridis(option = "magma") +
scale_color_viridis(option = "magma")
```

>>>>

---
title: "Reading information from the American Community Survey into R"
source: new
date: "2021-09-29"
categories:
- Recommendation
tags:
- Incomplete page
- Rmarkdown graphs
- Secondary data
output: html_document
page_update: complete
---

@@

```{r read-acs-data-01}
library(tidycensus)
library(tidyverse)
library(viridis)
key <- Sys.getenv("CENSUS_API_KEY")
census_api_key("11bafd82737f7d75b3305ada1e772501d0eab002")
tarr <- get_acs(geography = "tract", variables = "B19013_001",
state = "TX", county = "Tarrant", geometry = TRUE)
ggplot(tarr, aes(fill = estimate, color = estimate)) +
geom_sf() +
coord_sf(crs = 26914) +
scale_fill_viridis(option = "magma") +
scale_color_viridis(option = "magma")
```

### 2024-07-06 text/21/venn-diagram.Rmd
Updated 2022 Rmd files

```{r setup, echo=FALSE}
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(magrittr)))
```

```{r, fig.width=5, fig.height=5, echo=FALSE}
o <- 0.6
v <- 0.85
draw_circles <- function() {
  dx <- 1
  dy <- 0.866
  i <- seq(0, 2*pi, length=100)
  circle <- data.frame(x=sin(i), y=cos(i))
  circle %>%
    mutate(g=1) %>%
    mutate(x=x-o*dx) %>%
    mutate(y=y+o*dy) -> circle1
  circle %>%
    mutate(g=2) %>%
    mutate(x=x+o*dx) %>%
    mutate(y=y+o*dy) -> circle2
  circle %>%
    mutate(g=3) %>%
    mutate(y=y-o*dy) -> circle3

@@

```{r, fig.width=5, fig.height=5, echo=FALSE}
o <- 0.88
draw_circles()
```

>>>>

```{r venn-diagrams-01, echo=FALSE}
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(magrittr)))
```

```{r venn-diagrams-02, fig.width=5, fig.height=5, echo=FALSE}
o <- 0.6
v <- 0.85
draw_circles <- function() {
  dx <- 1
  dy <- 0.866
  i <- seq(0, 2*pi, length=100)
  circle <- data.frame(x=sin(i), y=cos(i))
  circle %>%
    mutate(g=1) %>%
    mutate(x=x-o*dx) %>%
    mutate(y=y+o*dy) -> circle1
  circle %>%
    mutate(g=2) %>%
    mutate(x=x+o*dx) %>%
    mutate(y=y+o*dy) -> circle2
  circle %>%
    mutate(g=3) %>%
    mutate(y=y-o*dy) -> circle3

@@

```{r venn-diagrams-03, fig.width=5, fig.height=5, echo=FALSE}
o <- 0.88
draw_circles()
```

### 2024-07-06 text/21/very-simple-splines.Rmd
Updated 2022 Rmd files

```{r}
library(ggplot2)
x <- 1:20
y <- c(rep(40, 15), 40-(1:5)^2)+round(rnorm(10), 1)
simulated_example <- data.frame(x, y)
simulated_example
ggplot(simulated_example, aes(x, y)) +
  geom_point()
```

@@

```{r}
linear_spline_ivs <- data.frame(x0=rep(1, 20), x1=c(rep(0, 15), 1:5))
linear_spline_ivs
```

@@

```{r}
simple_spline_data <- 
  data.frame(
    x1=linear_spline_ivs$x1, 
    x=simulated_example$x,
    y=simulated_example$y)
linear_spline_model <- lm(y~x1, data=simple_spline_data)
tidy(linear_spline_model)
```

@@

```{r}
x_extra <- seq(0, 20, length=1000)
data.frame(
  x=x_extra,
  x1=pmax(0, x_extra-15),
  x1sq=pmax(0, x_extra-15)^2) %>%
  full_join(simulated_example, by="x") -> enhanced_data
```

```{r}
linear_spline_predictions <- augment(linear_spline_model, newdata=enhanced_data)
ggplot(linear_spline_predictions, aes(x, y)) +
  geom_point() +
  geom_line(y=linear_spline_predictions$.fitted)
```

@@

```{r}
simple_spline_data %>%
  mutate(x1sq=x1*x1) -> quadratic_spline_data
quadratic_spline_model <- lm(y~x1sq, data=quadratic_spline_data)
tidy(quadratic_spline_model)
quadratic_spline_predictions <- augment(quadratic_spline_model, newdata=enhanced_data)
ggplot(quadratic_spline_predictions, aes(x, y)) +
  geom_point() +
  geom_line(y=quadratic_spline_predictions$.fitted)
```

@@

```{r}
x <- seq(0, 1, length=100)
y <- 3*x^2-2*x^3
data.frame(x=x, y=y) %>%
  ggplot(aes(x, y)) +
    geom_line()
```

@@

```{r}
x <- seq(0, 1, length=100)
y <- 3*x^2-2*x^3
data.frame(x=x, y=y) %>%
  mutate(y=0.75-0.5*y) %>%
  ggplot(aes(x, y)) +
    geom_line() + 
    expand_limits(y=0:1)
```

@@

```{r}
par(mar=rep(0.1,4))
plot(c(-1,2), c(0, 216), type="n", axes=FALSE, xlab="", ylab="")

>>>>

```{r very-simple-splines-01}
library(ggplot2)
x <- 1:20
y <- c(rep(40, 15), 40-(1:5)^2)+round(rnorm(10), 1)
simulated_example <- data.frame(x, y)
simulated_example
ggplot(simulated_example, aes(x, y)) +
  geom_point()
```

@@

```{r very-simple-splines-02}
linear_spline_ivs <- data.frame(x0=rep(1, 20), x1=c(rep(0, 15), 1:5))
linear_spline_ivs
```

@@

```{r very-simple-splines-03}
simple_spline_data <- 
  data.frame(
    x1=linear_spline_ivs$x1, 
    x=simulated_example$x,
    y=simulated_example$y)
linear_spline_model <- lm(y~x1, data=simple_spline_data)
tidy(linear_spline_model)
```

@@

```{r very-simple-splines-04}
x_extra <- seq(0, 20, length=1000)
data.frame(
  x=x_extra,
  x1=pmax(0, x_extra-15),
  x1sq=pmax(0, x_extra-15)^2) %>%
  full_join(simulated_example, by="x") -> enhanced_data
```

```{r very-simple-splines-05}
linear_spline_predictions <- augment(linear_spline_model, newdata=enhanced_data)
ggplot(linear_spline_predictions, aes(x, y)) +
  geom_point() +
  geom_line(y=linear_spline_predictions$.fitted)
```

@@

```{r very-simple-splines-06}
simple_spline_data %>%
  mutate(x1sq=x1*x1) -> quadratic_spline_data
quadratic_spline_model <- lm(y~x1sq, data=quadratic_spline_data)
tidy(quadratic_spline_model)
quadratic_spline_predictions <- augment(quadratic_spline_model, newdata=enhanced_data)
ggplot(quadratic_spline_predictions, aes(x, y)) +
  geom_point() +
  geom_line(y=quadratic_spline_predictions$.fitted)
```

@@

```{r very-simple-splines-07}
x <- seq(0, 1, length=100)
y <- 3*x^2-2*x^3
data.frame(x=x, y=y) %>%
  ggplot(aes(x, y)) +
    geom_line()
```

@@

```{r very-simple-splines-08}
x <- seq(0, 1, length=100)
y <- 3*x^2-2*x^3
data.frame(x=x, y=y) %>%
  mutate(y=0.75-0.5*y) %>%
  ggplot(aes(x, y)) +
    geom_line() + 
    expand_limits(y=0:1)
```

@@

```{r very-simple-splines-09}
par(mar=rep(0.1,4))
plot(c(-1,2), c(0, 216), type="n", axes=FALSE, xlab="", ylab="")
